Title

### 2014-06-05 TCAD2014/TCAD14.tex
add case study

\usepackage{multirow}

@@

Online arithmetic has been widely used in numerous applications such as signal processing and control algorithms [xxx]. Online arithmetic was originally designed for digit serial operation, as illustrated in Fig xx. It can be seen that in order to generate the first output digit, $\delta$ digits of inputs are required, where $\delta$ is called the ``online delay''. Normally $\delta$ a small constant, which is independent of the precision. For ease of discussion, for the rest of this paper, the input data is assumed to be fixed point numbers in the range $(-1,1)$. Based on this premise, the online representation of $N$-digit operands and result at iteration $j$ are given by (xx), where $j\in[-\delta,N-1]$ and $r$ denotes the radix [xxx].
\begin{eqnarray}\label{Eq:Online_Operands}
\footnotesize
  X_{[j]}=\sum_{i=1}^{j+\delta}x_ir^{-i},~Y_{[j]}=\sum_{i=1}^{j+\delta}y_ir^{-i},~Z_{[j]}=\sum_{i=1}^{j}z_ir^{-i}
\normalsize
\end{eqnarray}

MSD first operation is possible only if a redundant number system is used. Normally there are two most commonly used redundant number representations: carry-save (CS) [xxx] and signed-digit (SD) [xx]. With SD representation, each digit is represented using a redundant digit set $\{-a,\cdots,-1,0,1,\cdots,a\}$ where $a\in[r/2,r-1]$. In comparison, the standard non-redundant representation only uses a digit set $\{0,\cdots, r−1\}$. Thus a standard number corresponds to several possible redundant representations. For example, the binary number $0.011$ can be represented in SD form as 0.111, 0.101 or 0.011 among many other possible representations.
Due to the redundancy, the MSDs of the result can be calculated using partial information from both inputs. Then the value of the number can be revised using the subsequent digits, because each number has multiple representations.

\subsection{Binary Online Addition}

Adders serve as a critical building block for arithmetic operations. To perform digit-parallel online addition, a redundant adder can be used directly. The structure of an online adder where all signals represented with SD numbers of digit set $\{−1,0,1\}$ is shown in Fig. xxx. The module ``3:2'' denotes a 3:2 compressor, which takes three inputs and generates two outputs, and is logically equivalent to a full adder (FA). A major advantage of the redundant number system over the standard ripple-carry based arithmetic is that the propagation of carry is eliminated, resulting in a precision-independent computation time for addition. As labelled in Fig. xxx, ideally the computation delay of this adder is only two FA delays for any operand word-length, with the cost of one extra FA for each digit of operands. This makes the online adder suitable for building up more complex arithmetic operators such as multipliers to accelerate the sum of partial products [xx].

\subsection{Binary Online Multiplication}

Multiplication is another key arithmetic operator. Typically, online multiplication is performed in a recursive digit-serial manner, as illustrated in Algorithm \ref{Algorithm:OnlineMult} where both inputs and outputs are $N$-digit number as represented in~(\ref{Eq:Online_Operands}). For a given iteration $j$, the product digit $z_j$ is generated through a selection function $sel()$. For the radix $r$ and a chosen digit set, there exits an appropriate selection method and a value of $\delta$ which ensure convergence. As radix-2 is used most commonly in computer arithmetic, we keep $r=2$ throughout this paper with the corresponding redundant digit set $\{\overline{1},0,1\}$. In this case we have $\delta=3$, and $sel()$ is given by (\ref{Eq:SelFunc_OM}), from which it can be seen that the decision is made only based on 1 integer bit and 1 fractional bit of $W_{[j]}$.

\begin{algorithm}{\bf Online Multiplication} \label{Algorithm:OnlineMult}
  \begin{algorithmic}[1]
 \State  {\bf Initialization:}   $X_{[-\delta]}=Y_{[-\delta]}=P_{[-\delta]}=0$
  \State  {\bf Recurrence:}   $for~~ j=-\delta,~-\delta+1,~\cdots,~N-1 ~~do$
 \State  {\bf Compute:}    \label{Eq:OnlineMult_General}
\State  $    H_{[j]}  = r^{-\delta}\left(x_{j+\delta+1}\cdot Y_{[j+1]}+y_{j+\delta+1}\cdot X_{[j]}\right)$\label{Eq:OnlineMult_H}
      \State   $  W_{[j]}    = P_{[j]} + H_{[j]}$
     \State    $  z_j        =  sel(W_{[j]})$
    \State     $  P_{[j+1]}  =  r\left(W_{[j]}-z_j\right)$\label{Eq:OnlineMult_P}
  \end{algorithmic}
\end{algorithm}

\begin{equation}\label{Eq:SelFunc_OM}
\begin{array}{ll}
  sel(W_{[j]})=
    1 & if \ W_{[j]} \geq \frac{1}{2} \\
    0 &  if  \ -\frac{1}{2}\leq W_{[j]}<\frac{1}{2} \\
    \overline{1} & if \ W_{[j]}<-\frac{1}{2}
\end{array}
\end{equation}

\begin{figure}[h!]
\begin{center}
\includegraphics[width=7cm,height=4.5cm]{./Figures/circuit}
\caption{  . . . . . . . . . . . . }\label{circuit}
\end{center}
\end{figure}

..........................................\\

Let $\mu_{OM}$ denote the worst-case delay of an OM. It follows that if the clock period $T_S$ is greater than $\mu_{OM}$, correct results will be sampled. If, however, faster-than-rated sampling frequencies are applied such that $T_S<\mu_{OM}$, timing violations might happen and intermediate results will be sampled, potentially generating errors. For a given $T_S$, the maximum length of error-free propagation is described by (\ref{Eq:MaxChainLength}) where $f_S$ denotes the sampling frequency. We always ensure that the first digit of the product will be generated correctly, i.e. $b>\delta$.
%
\begin{eqnarray}\label{Eq:MaxChainLength}
\footnotesize
  b:=\left\lceil \frac{T_S}{\mu} \right\rceil=\left\lceil \frac{1}{\mu\cdot f_S}\right\rceil
  %b:=\left\lceil T_S/{\mu} \right\rceil=\left\lceil {1}/{(\mu\cdot f_S)}\right\rceil
\normalsize
\end{eqnarray}
%
We now determine when this timing constraint is not met and the size of error in this case.\\

\section{Probability of Timing Violations}\label{section:Prob}

@@

\begin{equation}\label{Eq:Bound_GenChain}
 -\delta\leq \tau\leq N-1-b=\tau_{max}
\end{equation}

\begin{equation}\label{Eq:Bound_ChainLength}
  b<d(\tau)\leq N-1-\tau
\end{equation}
%

\noindent However, the actual length of a ``carry'' chain is dependent upon input patterns. We then examine the relationship between $d(\tau)$ and the inputs that corresponds to the generation, propagation and annihilation of this carry chain.  Let the specific input pattern of stage $S_{\tau}$ be represented by $C(\tau)$, which can be classified into four types, as listed in (\ref{Eq:class}).

@@

\noindent The classification is based on whether a digit is zero, as all internal signals are reset to zero initially. Under the assumption that all digits are mutually independent and uniformly sampled from the digit set  $\{\overline{1},0,1\}$ as given in Section~\ref{Sec:Background}, the probabilities of $C_1,...,C_4$ are given in (\ref{Eq:probcaser}).

@@

Substituting (\ref{Eq:probcaser}) into \ref{Eq:PChainRule_Exp} yields (\ref{Eq:PChainRule_raw}).

\begin{equation}\label{Eq:PChainRule_raw}
\scriptsize
Pr_{\tau,d(\tau )}= \left(\frac{1}{(2a+1)^2}\right)^{\tau -1} \left(\frac{(2a+1)^2-1}{(2a+1)^2}\right)^{d(\tau )} \frac{1}{(2a+1)^2}
\normalsize
\end{equation}

@@

For all possible values of $\tau$ and $d(\tau)$, the possibility that timing violaions happen is given by $Pr$ in (\ref{Eq:Prob_timingViol}). According to (\ref{Eq:Bound_GenChain}) and (\ref{Eq:Bound_ChainLength}), the ranges in which the parameters $\tau$ and $d(\tau)$ are defined depend upon the value of the swamping period $T_S$. Hence $Pr$ is a function of $T_S$.

@@

\section{Error Estimation}

@@

\begin{figure}[tbp]
\begin{center}
\includegraphics[width=8cm,height=5cm]{./Figures/Error_Function_Spline_experiment}
\caption{Variances of overclocking errors. The results are obtained from the proposed models (blue line) and the experiments (red stars). (Legend and xylabels to be added.) }\label{Error_Function_Spline_experiment}
\end{center}
\end{figure}

%\noindent The {\bf Expectation of Overclocking Error} is also function of the sampling period and it is obtained by combining the values of $Pr$ and $\varepsilon$ as in (\ref{eovc}).

%\begin{equation}\label{eovc}
%E_{ovc}(T_S)= Pr(T_S) \cdot  v(T_S)
%\end{equation}

@@

% You can push biographies down or up by placing
% a \vfill before or after them. The appropriate
% use of \vfill depends on what kind of text is
% on the last page and whether or not the columns
% are being equalized.

>>>>

\usepackage{multirow}
\usepackage{subfigure}

@@

Online arithmetic has been widely used in numerous applications such as signal processing and control algorithms~\cite{Online_FPGADSP,Online_Control}. Online arithmetic was originally designed for digit serial operation, as illustrated in Fig.~\ref{Fig:OnlineDataFlow}. It can be seen that in order to generate the first output digit, $\delta$ digits of inputs are required, where $\delta$ is called the ``online delay''. Normally $\delta$ a small constant, which is independent of the precision. For ease of discussion, for the rest of this paper, the input data is assumed to be fixed point numbers in the range $(-1,1)$. Based on this premise, the online representation of $N$-digit operands and result at iteration $j$ are given by (\ref{Eq:Online_Operands}), where $j\in[-\delta,N-1]$ and $r$ denotes the radix~\cite{Ercegovac_Book}.

\begin{figure}[tbp]
  \centering
  \includegraphics[width=.48\textwidth]{./Figures/OnlineArithmetic_DataFlow.pdf}
  %\vspace{-4ex}
  \caption{Dataflow in digit-serial online arithmetic, in which both inputs and outputs are processed from the MSD to the LSD. $\delta$ denotes the online delay.}
  %\vspace{-1ex}
  \label{Fig:OnlineDataFlow}
\end{figure}
%
\begin{eqnarray}\label{Eq:Online_Operands}
\footnotesize
  X_{[j]}=\sum_{i=1}^{j+\delta}x_ir^{-i},~Y_{[j]}=\sum_{i=1}^{j+\delta}y_ir^{-i},~Z_{[j]}=\sum_{i=1}^{j}z_ir^{-i}
\normalsize
\end{eqnarray}

MSD first operation is possible only if a redundant number system is used. Normally there are two most commonly used redundant number representations: carry-save (CS)~\cite{CSadder}  and signed-digit (SD)~\cite{RedundantNumber}. With SD representation, each digit is represented using a redundant digit set $\{-a,\cdots,-1,0,1,\cdots,a\}$ where $a\in[r/2,r-1]$. In comparison, the standard non-redundant representation only uses a digit set $\{0,\cdots,r-1\}$. Thus a standard number corresponds to several possible redundant representations. For example, the binary number $0.011$ can be represented in SD form as $0.1\overline{1}1$, $0.10\overline{1}$ or $0.011$ among many other possible representations.

Due to the redundancy, the MSDs of the result can be calculated using partial information from both inputs. Then the value of the number can be revised using the subsequent digits, because each number has multiple representations.

\subsection{Binary Online Addition}\label{subsec:OnlineAdder}

Adders serve as a critical building block for arithmetic operations. To perform digit-parallel online addition, a redundant adder can be used directly. The structure of an online adder where all signals represented with SD numbers of digit set $\{-1,0,1\}$ is shown in Fig.~\ref{Fig:Radix2SD_adder}. The module ``3:2'' denotes a 3:2 compressor, which takes three inputs and generates two outputs, and is logically equivalent to a full adder (FA). A major advantage of the redundant number system over the standard ripple-carry based arithmetic is that the propagation of carry is eliminated, resulting in a precision-independent computation time for addition. As labelled in Fig.~\ref{Fig:Radix2SD_adder}, ideally the computation delay of this adder is only two FA delays for any operand word-length, with the cost of one extra FA for each digit of operands. This makes the online adder suitable for building up more complex arithmetic operators such as multipliers to accelerate the sum of partial products~\cite{RedundantMult_1987}.

\begin{figure}[tbp]
  \centering
  \includegraphics[width=.5\textwidth]{./Figures/SDAdder2.pdf}
  \vspace{-6ex}
  \caption{An $N$-digit binary digit-parallel online adder. Both inputs and outputs are represented using SD representation. ``3:2'' denotes a 3:2 compressor.}
    %\vspace{2ex}
  \label{Fig:Radix2SD_adder}
\end{figure}

\subsection{Binary Online Multiplication}\label{subsec:OnlineMultiplier}

Multiplication is another key arithmetic operator. Typically, online multiplication is performed in a recursive digit-serial manner, as illustrated in Algorithm \ref{Algorithm:OnlineMult}~\cite{Ercegovac_OnlineMult} where both inputs and outputs are $N$-digit number as represented in~(\ref{Eq:Online_Operands}). For a given iteration $j$, the product digit $z_j$ is generated through a selection function $sel()$. For the radix $r$ and a chosen digit set, there exits an appropriate selection method and a value of $\delta$ which ensure convergence. As the binary radix is used most commonly in computer arithmetic, we keep $r=2$ throughout this paper with the corresponding redundant digit set $\{\overline{1},0,1\}$. In this case $sel()$ is given by (\ref{Eq:SelFunc_OM}), and the selection is based on two integer digits and one fractional digit of $W_{[j]}$~\cite{Ercegovac_OnlineMult}.

\begin{algorithm}[tbp]
  \caption{Online Multiplication}  \label{Algorithm:OnlineMult}
  \begin{algorithmic}[1]
   \State \textbf{Initialization:}~$X_{[-\delta]}=Y_{[-\delta]}=P_{[-\delta]}=0$\vspace{.5ex}
    \For{$j=-\delta,~-\delta+1,~\cdots,~2N-1$}\vspace{.5ex}
        \State $H_{[j]} \leftarrow r^{-\delta}\left(x_{j+\delta+1}\cdot Y_{[j+1]}+y_{j+\delta+1}\cdot X_{[j]}\right)$\vspace{.5ex}
        \State $W_{[j]} \leftarrow P_{[j]} + H_{[j]}$\vspace{.5ex}
        \State $z_{j}~ \leftarrow sel(W_{[j]})$\vspace{.5ex}
        \State $P_{[j+1]} \leftarrow r\left(W_{[j]}-Z_{[j]}\right)$\vspace{.5ex}
    \EndFor
  \end{algorithmic}
\end{algorithm}

\begin{eqnarray}\label{Eq:SelFunc_OM}
%\small
  sel(W_{[j]})=\begin{cases}
    1 & \text{ if } W_{[j]} \geqslant \frac{1}{2} \\
    0 & \text{ if } -\frac{1}{2}\leqslant W_{[j]}<\frac{1}{2} \\
    \overline{1} & \text{ if } W_{[j]}<-\frac{1}{2}
  \end{cases}
%\normalsize
\end{eqnarray}

Algorithm \ref{Algorithm:OnlineMult} can be synthesized into a unrolled digit parallel structure as shown in  Figure~\ref{Fig:Radix2OnlineMultiplier}(a). In an $N$-digit online multiplier (OM), there are totally $N+\delta$ stages, of which the online inputs are generated from the appending logic according to (\ref{Eq:Online_Operands}). In each stage, an online adder is used as shown in Figure~\ref{Fig:Radix2OnlineMultiplier}(b). The SDVM module performs the signed-digit vector multiplication. When $r=2$ and the digit set $\{\overline{1},0,1\}$ is used, the implementation of SDVM is straight forward, because for instance the output of $(y_{j+\delta+1}\cdot X_{[j]})$ is 0, $X_{[j]}$ or $-X_{[j]}$ when $y$ is equal to 0, 1 and $-1$ respectively.

\begin{figure}[tbp]
  \centering
  \includegraphics[width=.49\textwidth]{./Figures/OM_Unrolled.pdf}
  %$\vspace{-4ex}
  \caption{(a) Synthesis of Algorithm 1 into a digit-parallel online multiplier (b) Structure of one stage, in which the maximum digit widths of signals are labeled.}
  %\vspace{-2ex}
  \label{Fig:Radix2OnlineMultiplier}
\end{figure}

Instead of duplicating the digit serial implementation $(N+\delta)$ times, each stage in the digit parallel architecture can be optimized for area reduction. For instance, the SDVM modules and the appending logic are not required in the last $\delta$ stages because the inputs are 0. This leads to a smaller online adder in these stages. Similarly for the first $\delta$ stages the selection logic can be removed, as the first digit of the result is generated at stage 0 ($S_0$).

\section{Probability Model of Overclocking Error}

As it is unlikely that timing violation happens in the online adder, in this Section we model the overclocking error in the radix-2 OM. From Figure~\ref{Fig:Radix2OnlineMultiplier} we observe two types of delay chains. One is caused by generation and propagation of $P_{[j]}$ among different stages. The other is the generation of online inputs $X_{[j]}$ and $Y_{[j]}$ from the appending logic. Since the appending logic is basically wires and simple combinational logic \cite{Online_Conversion}, the overall latency will eventually be determined by the delay of the $P_{[j]}$ path, especially with increasing operand word-lengths. As such, we initially model the delay of each stage within an OM to be a constant value $\mu$, as shown in Figure~\ref{Fig:Radix2OnlineMultiplier}(b). We also assume that the generation of online inputs costs no delay.
%\vspace{-1ex}

Let $\mu_{OM}$ denote the worst-case delay of an OM. It follows that if the clock period $T_S$ is greater than $\mu_{OM}$, correct results will be sampled. If, however, faster-than-rated sampling frequencies are applied such that $T_S<\mu_{OM}$, timing violations might happen and intermediate results will be sampled, potentially generating errors. For a given $T_S$, the maximum length of error-free propagation is described by (\ref{Eq:MaxChainLength}) where $f_S$ denotes the sampling frequency. We always ensure that the first digit of the product will be generated correctly, i.e. $b>\delta$.
%
\begin{eqnarray}\label{Eq:MaxChainLength}
\footnotesize
  b:=\left\lceil \frac{T_S}{\mu} \right\rceil=\left\lceil \frac{1}{\mu\cdot f_S}\right\rceil
  %b:=\left\lceil T_S/{\mu} \right\rceil=\left\lceil {1}/{(\mu\cdot f_S)}\right\rceil
\normalsize
\end{eqnarray}
%
We now determine when this timing constraint is not met and the size of error in this case.

\subsection{Probability of Timing Violations}\label{subsection:Prob}

@@

\begin{equation}\label{Eq:Bound_GenChain}
 -\delta\leq \tau\leq N-1-b=\tau_{max}
\end{equation}
\vspace{-3ex}
%
\begin{equation}\label{Eq:Bound_ChainLength}
  b<d(\tau)\leq N-1-\tau
\end{equation}
%

However, the actual length of a ``carry'' chain is dependent upon input patterns. We then examine the relationship between $d(\tau)$ and the inputs that corresponds to the generation, propagation and annihilation of this carry chain.  Let the specific input pattern of stage $S_{\tau}$ be represented by $C(\tau)$, which can be classified into four types, as listed in (\ref{Eq:class}).

@@

The classification is based on whether a digit is zero, as all internal signals are reset to zero initially. Under the assumption that all digits are mutually independent and uniformly sampled from the digit set  $\{\overline{1},0,1\}$ as given in Section~\ref{Sec:Background}, the probabilities of $C_1,...,C_4$ are given in (\ref{Eq:probcaser}).

@@

Substituting (\ref{Eq:probcaser}) into (\ref{Eq:PChainRule_Exp}) yields (\ref{Eq:PChainRule_raw}).

\begin{equation}\label{Eq:PChainRule_raw}
\footnotesize
  Pr_{\tau,d(\tau )}= \left(\frac{1}{(2a+1)^2}\right)^{\tau -1} \left(\frac{(2a+1)^2-1}{(2a+1)^2}\right)^{d(\tau )} \frac{1}{(2a+1)^2}
\normalsize
\end{equation}

@@

For all possible values of $\tau$ and $d(\tau)$, the possibility that timing violations happen is given by $Pr$ in (\ref{Eq:Prob_timingViol}). According to (\ref{Eq:Bound_GenChain}) and (\ref{Eq:Bound_ChainLength}), the ranges in which the parameters $\tau$ and $d(\tau)$ are defined depend upon the value of the swamping period $T_S$. Hence $Pr$ is a function of $T_S$.

@@

\subsection{Estimation of Timing Errors}

@@

\begin{figure}[tbp]
\centering
  \includegraphics[width=8cm,height=5cm]{./Figures/Error_Function_Spline_experiment}
  \caption{Variances of overclocking errors. The results are obtained from the proposed models (blue line) and the experiments (red stars). (Legend and xylabels to be added.) }\label{Error_Function_Spline_experiment}
\end{figure}

\section{Case Study: Image Filter}\label{Sec:CaseStudy}
\subsection{Experimental Setup}
The benefits of the proposed methodology are demonstrated by using a $3\times3$ Gaussian image filter, which is implemented with two types of computer arithmetic. One is the binary online arithmetic while the other is the binary traditional arithmetic. For design with online arithmetic, all inputs and outputs are represented in the online format as given in (\ref{Eq:Online_Operands}). The basic building blocks: adders and multipliers as described in Section~\ref{subsec:OnlineAdder} and Section~\ref{subsec:OnlineMultiplier} are employed. For design with traditional arithmetic, all signals are represented using the 2's complement format. We build this type of image filter by using speed-optimized adders and multipliers which are created using Xilinx Core Generator \cite{XilinxMult}.

In order to achieve the desired latency between input and output, both designs are overclocked and the errors seen at the output are recorded. The results are obtained from a Xilinx Virtex-6 FPGA xxx through post place and route simulations. The results are evaluated in terms of mean relative error (MRE), which represents the percentage of error at outputs, as given by (\ref{Eq:MRE}) where $E_{error}$ and $E_{out}$ refer to the mean value of error and correct output, respectively.
%
\begin{eqnarray}\label{Eq:MRE}
\scriptsize
  MRE=\left|\frac{E_{error}}{E_{out}}\right|\times100\%
  %MRE=\left|{E_{error}}/{E_{out}}\right|\times100\%
\end{eqnarray}
\normalsize

In our experiments, two types of input data are utilized. One is randomly sampled from a uniform distribution of $N$-digit numbers. This type is referred to as ``Uniform Independent (UI) inputs''. The other is called ``real inputs'', which are the pixel values of several $512\times512$ benchmark images.

\subsection{Quantify the Impact of Overclocking}
The MRE values of the image filter with traditional arithmetic (dotted lines) and online arithmetic (solid lines) when $N=8$ are illustrated in Fig.~\ref{Fig:MRE_ImageFilter}. According to the timing analysis tool, the rated operating frequencies of the two designs are $169$MHz and $148$MHz, respectively. However as seen in Fig.~\ref{Fig:MRE_ImageFilter}, for the UI inputs, design with online arithmetic actually operates at a higher frequency without timing violations in comparison to the design using traditional arithmetic. This error-free frequency is even larger when using the real image data as inputs, since the real data do not exactly follow the uniform distribution or the independent assumption. In Fig.~\ref{Fig:MRE_ImageFilter} the ``Lena'' benchmark image is used as the real inputs.

\begin{figure}
    \centering
    \includegraphics[width=.48\textwidth]{./Figures/Exps/MRE.eps}
    %\vspace{-4ex}
    \caption{Overclocking error in an image filter with two types of computer arithmetic: online arithmetic and standard binary arithmetic, of which the rated frequencies are 148.3MHz and 168.7MHz, respectively, according to the timing analysis tool.}
    \label{Fig:MRE_ImageFilter}
    %\vspace{-3ex}
\end{figure}

If errors can be tolerated, we may allow timing violations to happen for better performance. The sensitivity of overclocking for a given arithmetic can be evaluated by the data slope in Figure~\ref{Fig:MRE_ImageFilter}. For instance under an error budget of $1\%$~MRE, using UI inputs the frequency speedup of the traditional design is $3.89\%$ with respect to the maximum frequency without errors, whereas the design with online arithmetic can be overclocked by $6.85\%$. This indicates that online arithmetic is less sensitive to overclocking, as discussed in Section~\ref{Sec:MeanError}. The difference is greater using real image data: $13.74\%$ frequency speedup with online arithmetic against $4.04\%$ with traditional arithmetic, because longer chains happen with a smaller probability with real inputs.
%\vspace{-1ex}

The output images for both design scenarios are presented in Fig.~\ref{Fig:LenaImage}. Since the overclocking errors are in the least significant ends of the results with online arithmetic, the degradation on the image can be hardly observed. In contrast, timing violations cause error in the MSDs with traditional arithmetic. This leads to ``salt and pepper noise'' and severe quality loss as shown on the images in the right column of Fig.~\ref{Fig:LenaImage}. Furthermore, errors in the MSDs result in large noise power. Hence the signal-to-noise ratio (SNR) of the traditional design is small.

\begin{figure}[htb]
  %\vspace{-2.5ex}
  \centering
  \subfigure[1.05$f_0$,~SNR=38.3dB]{
  \begin{minipage}[c]{0.24\textwidth}
    \centering
    \includegraphics[width=1.3in]{./Figures/Exps/Lena/Online_T404.eps}
  \end{minipage}%
  }% This is important! use % to indicate same line, otherwise new line
  \subfigure[1.05${f_0}'$,~SNR:21.5dB]{
  \begin{minipage}[c]{0.24\textwidth}
    \centering
    \includegraphics[width=1.3in]{./Figures/Exps/Lena/Trad_T426.eps}
  \end{minipage}
  }\vspace{-1ex}
  \subfigure[1.15$f_0$,~SNR:36.2dB]{
  \begin{minipage}[c]{0.24\textwidth}
    \centering
    \includegraphics[width=1.3in]{./Figures/Exps/Lena/Online_T370.eps}
  \end{minipage}
  }%
  \subfigure[1.15${f_0}'$,~SNR:9.0dB]{
  \begin{minipage}[c]{0.24\textwidth}
    \centering
    \includegraphics[width=1.3in]{./Figures/Exps/Lena/Trad_T390.eps}
  \end{minipage}
  }\vspace{-1ex}
  \subfigure[1.25$f_0$,~SNR:26.7dB]{
  \begin{minipage}[c]{0.24\textwidth}
    \centering
    \includegraphics[width=1.3in]{./Figures/Exps/Lena/Online_T340.eps}
  \end{minipage}%
  }% same line
  \subfigure[1.25${f_0}'$,~SNR:8.5dB]{
  \begin{minipage}[c]{0.24\textwidth}
    \centering
    \includegraphics[width=1.3in]{./Figures/Exps/Lena/Trad_T358.eps}
  \end{minipage}
  }
%\vspace{-2ex}
\caption{Output images of image filter using online arithmetic (left column) and traditional arithmetic (right column), where $f_0$ and ${f_0}'$ denote the maximum error-free frequencies for each design.}
\label{Fig:LenaImage}
%\vspace{-1ex}
\end{figure}

\subsection{Potential Benefits in Circuit Design}
In general, our results could be of interest to a circuit designers in two ways. By choosing different arithmetic and data representations, either a circuit can be designed to operate at a certain frequency with the minimum possible MRE, or a given error budget specified by the algorithm designer can be met with the fastest achievable frequency. For the first case, the experimental results obtained by UI inputs and 4 benchmark images are summarized in Table~\ref{Tab:MRE_Redu_Freq} in terms of the relative reduction of MRE as given by (\ref{Eq:MRE_Reduction}) where $MRE_{OL}$ and $MRE_{Trad}$ denote the value obtained with online arithmetic and with traditional arithmetic, respectively, and in Table \ref{Tab:SNR_Impro_Freq} for the differences of SNR.

We also perform experiments using other benchmark images. The results are summarized in Table~\ref{Tab:MRE_Redu_Freq} and Table~\ref{Tab:SNR_Impro_Freq} in terms of the relative reduction of MRE and the improvements of SNR, respectively. In both tables the frequency is normalized to the maximum error-free frequency for each arithmetic. From Table~\ref{Tab:MRE_Redu_Freq}, a significant reduction of MRE can be observed using online arithmetic. The geometric mean reduction of MRE is $89.2\%$ using UI data. Even larger reductions of MRE can be achieved when testing with real image data, varying from $97.3\%$ to $98.2\%$, as expected given the results shown in Figure~\ref{Fig:MRE_ImageFilter}. Similarly from Table~\ref{Tab:SNR_Impro_Freq}, the improvements in SNR are $21.4dB\sim43.9dB$.
%\vspace{-1ex}

\begin{eqnarray}\label{Eq:MRE_Reduction}
\scriptsize
  \frac{MRE_{Trad}-MRE_{OL}}{MRE_{Trad}}\times100\%
  %{(MRE_{Trad}-MRE_{OL})}/{MRE_{Trad}}\times100\%
\normalsize
\end{eqnarray}

\begin{table}[tbh]
\vspace{-1ex}
\renewcommand{\arraystretch}{1.1}
\setlength{\tabcolsep}{4.1pt}
\caption{Relative Reduction of MRE with Online Arithmetic for Various Normalized Frequencies.}
\vspace{1ex}
\label{Tab:MRE_Redu_Freq}
\small
%\scriptsize
\centering
\begin{tabular}{|c|ccccc|c|}
\hline
\multirow{2}*{\textbf{Inputs}} & \multicolumn{5}{c|}{\textbf{Normalized Frequency}} &
\multirow{2}*{\begin{tabular}{c}\textbf{Geo.}\\\textbf{Mean}\end{tabular}}\\
& 1.05 & 1.10 & 1.15 & 1.20 & 1.25 &\\
\hline
Uniform & 94.5\% & 89.1\% & 90.1\% & 88.3\% & 84.3\% & 89.2\%\\
Lena    & 99.3\% & 99.2\% & 98.9\% & 97.7\% & 94.9\% & 97.9\%\\
Pepper  & 99.7\% & 98.3\% & 98.1\% & 97.2\% & 95.1\% & 97.7\%\\
Sailboat& 99.5\% & 97.9\% & 97.3\% & 96.8\% & 95.1\% & 97.3\%\\
Tiffany & 99.9\% & 97.6\% & 98.4\% & 97.8\% & 97.2\% & 98.2\%\\
\hline
\end{tabular}
%\vspace{-1ex}
\normalsize
\end{table}

\begin{table}[thb]
%\vspace{-1ex}
\renewcommand{\arraystretch}{1.1}
\setlength{\tabcolsep}{4.1pt}
\caption{Improvement of SNR (dB) with Online Arithmetic for Various Normalized Frequencies.}
\vspace{1ex}
\label{Tab:SNR_Impro_Freq}
%\scriptsize
\small
\centering
\begin{tabular}{|c|ccccc|}
\hline
\multirow{2}*{\textbf{Inputs}} & \multicolumn{5}{c|}{\textbf{Normalized Frequency}} \\
& 1.05 & 1.10 & 1.15 & 1.20 & 1.25\\
\hline
Lena    & 44.6 & 36.3 & 33.2 & 29.1 & 22.9\\
Pepper  & 35.7 & 28.3 & 28.7 & 25.9 & 24.1\\
Sailboat& 33.7 & 27.5 & 26.3 & 25.0 & 21.7\\
Tiffany & 43.9 & 25.9 & 29.5 & 25.6 & 24.5\\
\hline
\end{tabular}
\vspace{-2ex}
\normalsize
\end{table}

For the second design perspective, Table~\ref{Tab:MRE_Reduc_Error} illustrates the frequency speedups with different input types when specific error budgets can be tolerated. We see that for all input types using online arithmetic still outperforms the traditional design for each MRE budget in terms of operating frequency. Likewise the geometric mean of frequency speed-ups is larger for real image inputs.\vspace{-1ex}

\begin{table}[tbh]
\vspace{-2ex}
\renewcommand{\arraystretch}{1.1}
\setlength{\tabcolsep}{4.1pt}
\caption{Relative Improvement in Frequency with Online Arithmetic for Various Error Budgets.}
\label{Tab:MRE_Reduc_Error}
\small
%\scriptsize
\centering
\begin{tabular}{|c|cccc|c|}
\hline
\multirow{2}*{\textbf{Inputs}} & \multicolumn{4}{c|}{\textbf{Error Budget}} &
\multirow{2}*{\begin{tabular}{c}\textbf{Geo.}\\\textbf{Mean}\end{tabular}}\\
& 0.01\% & 0.1\% & 1\% & 10\% &\\
\hline
Uniform & N/A & 4.59\%   & 8.78\%  & 12.83\% & 8.03\%\\
Lena    & 7.96\% & 11.50\%  & 16.39\% & 13.71\% & 11.88\%\\
Pepper  & 6.22\% & 8.87\%   & 18.68\% & 15.94\% & 11.32\%\\
Sailboat& 5.71\% & 8.87\%   & 16.93\% & 15.29\% & 10.70\%\\
Tiffany & 2.35\% & 6.90\%   & 18.58\% & 12.22\% & 7.79\% \\
\hline
\end{tabular}
\label{Tab:}
%\vspace{-2ex}
\normalsize
\end{table}

\subsection{Area Overhead of Our Approach}
The area comparison between two designs is illustrated in Table~\ref{Tab:AreaComparison}. While we notice that our approach comes at some increase in area, the FPGA architecture is optimized for conventional arithmetic. For instance, the Virtex series employs dedicated multiplexers and encoders for very fast ripple carry addition~\cite{Virtex6}. Besides, at least 2 orders of magnitude error reduction is obtained for a given frequency in our design, as shown in Figure \ref{Fig:MRE_ImageFilter}. Although the differences in both error and area can be compensated by using more digits with traditional arithmetic, this will result in longer delay and an even larger gap in frequency between two designs.

\begin{table}[tbh]
\vspace{-1ex}
\small
\caption{FPGA Resource Usage Comparison.}
\vspace{.5ex}
\centering
\label{Tab:AreaComparison}
\begin{tabular}{|c|cc|c|}
\hline
\multirow{2}*{\textbf{Metric}} & \multicolumn{2}{c|}{\textbf{Arithmetic Type}} &
\multirow{2}*{\textbf{Overhead}}\\
 & Traditional & Online &\\ \hline
 Look-Up Tables   & 912 & 1896 & 2.08 \\ \hline
 Slices & 324 & 525 & 1.62 \\ \hline

\end{tabular}
\vspace{-1ex}
\normalsize
\end{table}

@@
